Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-03-10 7.GWAS/7.GWAS-phase-transitions.tex
ch 7
i implemented my minor edits to ch 7.  I have 3 paragraphs in BLUE which need to be checked -- pages 97, 102, 105.

Recall Theorem \ref{thm:chi-squared-exact-approx-boundary} predicts that FWER and FNR can be simultaneously controlled in large dimensions if and only if 
\begin{equation}
    r = \frac{\lambda}{2\log{p}} = \frac{w^2n}{2\log{p}} > 1.
\end{equation}
Therefore, if we were to apply FWER-controlling procedures at low nominal levels (say, $5\%$), then the FNR would experience a phase transition in the sense that, if
\begin{equation} \label{eq:power-1-region}
    r>1 \iff w^2 > \frac{2\log{p}}{n},
\end{equation}
then the FNR can be close to 0; otherwise, FNR must be close to 1.

@@

\begin{example}
\label{exmp:OR-RAF_phase_transition}
Consider association tests on $2\times2$ contingency tables at $p$ locations as introduced in Section \ref{sec:motivation-chisq}, where the counts follow 
a multinomial distribution
% independent binomial distributions 
% $$
% O_{11} \sim \mathrm{Binom}(n\phi_1, fR/(fR+1-f)),\quad 
% O_{21} \sim \mathrm{Binom}(n(1-\phi_1), fR/(fR+1-f)),
% $$
parametrized by $(f, R, \phi_1)$ as in Section \ref{sec:optimal-design}.
Assume that the phenotype marginals are fixed at $\phi_1 = \phi_2 = 1/2$.
% --- as is the case in genetic association studies --- 
Applying Bonferroni's procedure with nominal FWER at $\alpha=5\%$ level, we can approximate the marginal power of association tests by
\begin{equation} \label{eq:power-approximation}
    \P[\chi^2_{1}(\lambda)>\chi^2_{1,\alpha/p}],
\end{equation}
where $\chi^2_{1,\alpha/p}$ is the upper $(\alpha/p)$-quantile of a central chi-squared distribution with 1 degree of freedom.
We calculate this marginal power as a function of the parameters $(f,R)$ in three scenarios:
\begin{itemize}
    \item $p=4$, $n=3\times10^4$ 
    \item $p=10^2$, $n=1\times10^5$
    \item $p=10^6$, $n=3\times10^6$
\end{itemize}
and visualize the results as heatmaps\footnote{Since genetic variants can always be relabelled such that Variant 1 is positively associated with Cases, we only produce part of the diagram where $R>1$.
Sample sizes marked in the figure are adjusted by a factor of $1/2$, to reflect the genetic context where a pair of alleles are measured for every individual at every genomic location.} (referred to as OR-RAF diagrams) in Figure \ref{fig:OR-RAF_GWAS}.
These parameter values are chosen so that $\log{p}/n$ are roughly constant (around $4.6\times10^{-5}$).

@@

If we wish to achieve exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}, Theorem \ref{thm:chi-squared-exact-approx-boundary} predicts that the signal size parameter $r$ has to be at least $\widetilde{g}(\beta)= 1$.
This signal size calls for a sample size of $n = \lambda / w^2 = 2r\log(p)/w^2 \approx 307,011$.
In a typically GWAS, a pair of alleles are sequenced for every marker location, bringing the required number of subjects in the study to $n/2 \approx 153,509$.
\end{example}

>>>>

Recall Theorem \ref{thm:chi-squared-exact-approx-boundary} predicts that FWER and FNR can be simultaneously controlled in large dimensions if and only if 
\begin{equation}
    r = \frac{\lambda}{2\log{p}} = \frac{w^2n}{2\log{p}} > 1.
\end{equation}
Therefore, if we were to apply FWER-controlling procedures at low nominal levels (say, $5\%$), then the FNR would experience a phase transition
in the following sense.  If the signal size is strong enough, i.e., 
\begin{equation} \label{eq:power-1-region}
    r>1 \iff w^2 > \frac{2\log{p}}{n},
\end{equation}
then the FNR can be close to 0; otherwise, FNR must be close to 1.

@@

\begin{example}
\label{exmp:OR-RAF_phase_transition}
Consider association tests on $2\times2$ contingency tables at $p$ locations as introduced in Section \ref{sec:motivation-chisq}, where the counts follow 
a multinomial distribution
% independent binomial distributions 
% $$
% O_{11} \sim \mathrm{Binom}(n\phi_1, fR/(fR+1-f)),\quad 
% O_{21} \sim \mathrm{Binom}(n(1-\phi_1), fR/(fR+1-f)),
% $$
parametrized by $(f, R, \phi_1)$ as in Section \ref{sec:optimal-design}.
Assume that the phenotype marginals are fixed at $\phi_1 = \phi_2 = 1/2$.
% --- as is the case in genetic association studies --- 
Applying Bonferroni's procedure with nominal FWER at $\alpha=5\%$ level, we can approximate the marginal power of association tests by
\begin{equation} \label{eq:power-approximation}
    \P[\chi^2_{1}(\lambda)>\chi^2_{1,\alpha/p}],
\end{equation}
where $\chi^2_{1,\alpha/p}$ is the upper $(\alpha/p)$-quantile of a central chi-squared distribution with 1 degree of freedom.
We calculate this marginal power as a function of the parameters $(f,R)$ in three scenarios:
\begin{itemize}
    \item $p=4$, $n=3\times10^4$ 
    \item $p=10^2$, $n=1\times10^5$
    \item $p=10^6$, $n=3\times10^6$
\end{itemize}
and visualize the results as heatmaps\footnote{Since genetic variants can always be relabelled such that Variant 1 is positively associated with Cases, we only produce part of the diagram where $R>1$.
Sample sizes marked in the figure are adjusted by a factor of $1/2$, to reflect the genetic context where a pair of alleles are measured for every individual at every genomic location.} (referred to as OR-RAF diagrams) in Figure \ref{fig:OR-RAF_GWAS}.
These parameter values are chosen so that $\log(p)/n$ are roughly constant (around $4.6\times10^{-5}$).

@@

If we wish to achieve exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}, Theorem \ref{thm:chi-squared-exact-approx-boundary} predicts that the signal size parameter $r$ has to be at least $\widetilde{g}(\beta)= 1$.
This signal size calls for a sample size of $n = \lambda / w^2 = 2r\log(p)/w^2 \approx 307,011$.
In a typical GWAS, a pair of alleles are sequenced for every marker location, bringing the required number of subjects in the study to $n/2 \approx 153,509$.
\end{example}

### 2021-03-10 7.GWAS/7.chi-square.tex
ch 7
i implemented my minor edits to ch 7.  I have 3 paragraphs in BLUE which need to be checked -- pages 97, 102, 105.

Similar to the analysis of additive error models in Chapter \ref{chap:phase-transitions}, we will work with triangular arrays of chi-square models \eqref{eq:model-chisq-Chapter6} indexed by $p$.
We adopt the same parametrization for the sparsity of the non-centrality parameter vectors $\lambda = \lambda_p$,
\begin{equation} \label{eq:signal-sparsity}
    |S_p| = \left\lfloor p^{1-\beta} \right\rfloor, \quad \beta\in(0,1]
\end{equation}
where $\beta$ parametrizes the problem sparsity.
The closer $\beta$ is to 1, the sparser the support $S_p$; conversely, when $\beta$ is close to 0, the support is dense with many non-null signals.

@@

The first main result characterizes the phase transition phenomenon in the exact support recovery problem under the chi-square model.

\begin{theorem} \label{thm:chi-squared-exact-boundary}
Consider the high-dimensional chi-squared model \eqref{eq:model-chisq-Chapter6} with signal sparsity and size as described in \eqref{eq:signal-sparsity} and \eqref{eq:signal-size}.
The function 
\begin{equation} \label{eq:exact-boundary-chisquared}
    g(\beta) = \left(1 + \sqrt{1-\beta}\right)^2
\end{equation}
characterizes the phase transition of exact support recovery problem.
Specifically, if $\underline{r} > {{g}}(\beta)$, then Bonferroni's, Sid\'ak's, Holm's, and Hochberg's procedures with slowly vanishing (see Definition \ref{def:slowly-vanishing}) nominal FWER levels all achieve asymptotically exact support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < {{g}}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, we have $\P[\widehat{S}_p=S_p]\to0$.
Therefore, in view of Lemma \ref{lemma:risk-exact-recovery-probability}, exact support recovery asymptotically fails for all thresholding procedures in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

The procedures listed in Theorem \ref{thm:chi-squared-exact-boundary} were reviewed in Section \ref{sec:statistical-procedures}. 
Proof of the theorem can be found in Section \ref{subsec:proof-chi-squared-exact-boundary}. 
% The boundary \eqref{eq:exact-boundary-chisquared} is plotted in Figure \ref{fig:phase-chi-squared}.

@@

\begin{remark} \label{rmk:strong-classification-boundary-2}
Theorem \ref{thm:chi-squared-exact-boundary} predicts that the asymptotic boundaries are the same for all values of the parameter $\nu$.
In simulations (Section \ref{sec:numerical}), we find this asymptotic prediction to be quite accurate for $\nu\le3$ even in moderate dimensions ($p=100$). 
For $\nu>3$, the phase transitions take place somewhat above the boundary ${g}$.
The behavior is qualitatively similar for the other three phase transitions (see Theorems \ref{thm:chi-squared-exact-approx-boundary}, \ref{thm:chi-squared-approx-boundary}, and \ref{thm:chi-squared-approx-exact-boundary} below).
\end{remark}

@@

The next theorem describes the phase transition in the exact-approximate support recovery problem.

\begin{theorem} \label{thm:chi-squared-exact-approx-boundary}
In the context of Theorem \ref{thm:chi-squared-exact-boundary}, 
the function 
\begin{equation} \label{eq:exact-approx-boundary-chisquared}
    \widetilde{g}(\beta) = 1
\end{equation}
characterizes the phase transition of exact-approximate support recovery problem.
Specifically, if $\underline{r} > \widetilde{g}(\beta)$, then the procedures listed in Theorem \ref{thm:chi-squared-exact-boundary} with slowly vanishing nominal FWER levels achieve asymptotically exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < \widetilde{g}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, the exact-approximate support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

Our third main result characterizes the phase transition phenomenon in the approximate support recovery problem in the chi-square model.

\begin{theorem} \label{thm:chi-squared-approx-boundary}
Consider the high-dimensional chi-squared model \eqref{eq:model-chisq-Chapter6} with signal sparsity and size as described in \eqref{eq:signal-sparsity} and \eqref{eq:signal-size}.
The function 
\begin{equation} \label{eq:approx-boundary-chisquared}
    h(\beta) = \beta
\end{equation}
characterizes the phase transition of approximate support recovery problem.
Specifically, if $\underline{r} > {h}(\beta)$, then the \ac{BH} procedure $\widehat{S}_p$ (defined in Section \ref{sec:statistical-procedures}) with slowly vanishing (see Definition \ref{def:slowly-vanishing}) nominal FDR levels achieves asymptotically approximate support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < {h}(\beta)$, then approximate support recovery asymptotically fails in the sense of \eqref{eq:support-recovery-failure} for all thresholding procedures.
\end{theorem}

@@

\begin{theorem} \label{thm:chi-squared-approx-exact-boundary}
In the context of Theorem \ref{thm:chi-squared-approx-boundary}, the function 
\begin{equation} \label{eq:approx-exact-boundary-chisquared}
    \widetilde{h}(\beta) = \left(\sqrt{\beta} + \sqrt{1-\beta}\right)^2
\end{equation}
characterizes the phase transition of approximate-exact support recovery problem.
Specifically, if $\underline{r} > \widetilde{h}(\beta)$, then the Benjamini-Hochberg procedure with slowly vanishing nominal FDR levels achieves asymptotically approximate-exact support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < \widetilde{h}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, the approximate-exact support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

Notice that all phase transitions boundaries are identical to those in the Gaussian additive error model \eqref{eq:model-additive} under one-side alternative.
We refer readers to Figure \ref{fig:phase-Gaussian-errors} in Section \ref{sec:additive-error-model-boundaries} for a visualization of the results in Theorems \ref{thm:chi-squared-exact-boundary} through \ref{thm:chi-squared-approx-exact-boundary}.

@@

The all four Theorems so far focus only on the idealized models \eqref{eq:model-chisq-Chapter6} where statistics are \emph{independent}.
Support recovery problems under dependent observations remain to be explored.
Recall in Chapter \ref{chap:phase-transitions} we showed that the boundary for the exact support recovery problem in the additive error model \eqref{eq:model-additive} continues to hold even under severe dependence and general distributional assumptions.
We conjecture that similar results would also hold, under classes of dependence structures that are ``not too different from independence'', in the chi-square models.
As an example, in the GWAS application, dependence among the genetic markers at different locations (known as linkage disequilibrium) decay as a function of their physical distances on the genome \citep{bush2012genome}, resulting in locally dependent test statistics.
It would be of great interest to extend the current theory to cover important dependence structures that arise in such applications.

@@

% $\mathrm{risk}^{\mathrm{EA}}$ and $\mathrm{risk}^{\mathrm{AE}}$.
As alluded to in Section \ref{sec:motivation-chisq} in the introduction, we draw explicit comparisons between the one-sided and two-sided alternatives in Gaussian additive error models \eqref{eq:model-additive}.
% The exact, and the approximate support recovery problems in the additive error model \eqref{eq:model-additive} under standard Gaussian errors have been studied in \cite{gao2018fundamental} and \cite{arias2017distribution}, respectively.

@@

Comparing Theorems \ref{thm:chi-squared-exact-approx-boundary} to \ref{thm:Gaussian-error-exact-approx-boundary} and Theorems \ref{thm:chi-squared-approx-exact-boundary} to \ref{thm:Gaussian-error-approx-exact-boundary}, we see that the phase transition boundaries under the two types of alternatives are also identical in the exact-approximate and approximate-exact support recovery problems.

@@

Therefore, all phase transition boundaries coincide with those in the additive error models obtained in Chapter \ref{chap:phase-transitions} under their respective parametrizations.
% and Figure \ref{fig:phase-chi-squared} continues to apply.
This indicates vanishing differences between the difficulties of the one-sided and two-sided alternatives in the Gaussian additive error model \eqref{eq:model-additive}.
The additional uncertainty in the two-sided alternatives do not call for larger signal sizes in these problems, asymptotically.

>>>>

Similar to the analysis of additive error models in Chapter \ref{chap:phase-transitions}, we will work with triangular arrays of chi-square models \eqref{eq:model-chisq-Chapter6} indexed by $p$.
We adopt the same parametrization for the sparsity of the non-centrality parameter vectors $\lambda = \lambda_p$,
\begin{equation} \label{eq:signal-sparsity}
    |S_p| = \left\lfloor p^{1-\beta} \right\rfloor, \quad \beta\in(0,1]
\end{equation}
where $S_p:= \{ i\, :\, \lambda(i) >0 \}$ is the {\em signal support} set and  $\beta$ parametrizes the problem {\em sparsity}.
The closer $\beta$ is to 1, the sparser the support $S_p$; conversely, when $\beta$ is close to 0, the support is dense with many non-null signals.

@@

The first main result characterizes the phase transition phenomenon in the exact support recovery problem under the chi-square model.  It parallels
Theorem \ref{thm:Gaussian-error-exact-boundary}.

\begin{theorem} \label{thm:chi-squared-exact-boundary}
Consider the high-dimensional chi-squared model \eqref{eq:model-chisq-Chapter6} with signal sparsity and size as described in \eqref{eq:signal-sparsity} and \eqref{eq:signal-size}.
The function 
\begin{equation} \label{eq:exact-boundary-chisquared}
    g(\beta) = \left(1 + \sqrt{1-\beta}\right)^2
\end{equation}
characterizes the phase transition of exact support recovery problem.
Namely, the following two results hold.\\

{\rm (i)} If $\underline{r} > {{g}}(\beta)$, then Bonferroni's, Sid\'ak's, Holm's, and Hochberg's procedures with slowly vanishing (see Definition \ref{def:slowly-vanishing}) nominal FWER levels all achieve asymptotically exact support recovery in the sense of \eqref{eq:support-recovery-success}.\\

{\rm (ii)} Conversely, if $\overline{r} < {{g}}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, we have $\P[\widehat{S}_p=S_p]\to0$.
Therefore, in view of Lemma \ref{lemma:risk-exact-recovery-probability}, exact support recovery asymptotically fails for all thresholding procedures in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

The procedures listed in Theorem \ref{thm:chi-squared-exact-boundary} were reviewed in Section \ref{sec:statistical-procedures}. 
The proof of the theorem can be found in Section \ref{subsec:proof-chi-squared-exact-boundary}. 
% The boundary \eqref{eq:exact-boundary-chisquared} is plotted in Figure \ref{fig:phase-chi-squared}.

@@

\begin{remark} \label{rmk:strong-classification-boundary-2}
Theorem \ref{thm:chi-squared-exact-boundary} predicts that the asymptotic boundaries are the same for all values of the degrees of freedom 
parameter $\nu$.  In simulations (Section \ref{sec:numerical}), we find this asymptotic prediction to be quite accurate for $\nu\le3$ even in 
moderate dimensions ($p=100$).  For $\nu>3$, the phase transitions take place somewhat above the boundary ${g}$.
The behavior is qualitatively similar for the other three phase transitions (see Theorems \ref{thm:chi-squared-exact-approx-boundary}, 
\ref{thm:chi-squared-approx-boundary}, and \ref{thm:chi-squared-approx-exact-boundary} below).
\end{remark}

@@

The next theorem describes the phase transition in the exact-approximate support recovery problem. Recall also
Theorem \ref{thm:Gaussian-error-exact-approx-boundary}.

\begin{theorem} \label{thm:chi-squared-exact-approx-boundary}
In the context of Theorem \ref{thm:chi-squared-exact-boundary}, 
the function 
\begin{equation} \label{eq:exact-approx-boundary-chisquared}
    \widetilde{g}(\beta) = 1
\end{equation}
characterizes the phase transition of exact-approximate support recovery problem.
 Namely, the following two results hold.\\

{\rm (i)} If $\underline{r} > \widetilde{g}(\beta)$, then the procedures listed in Theorem \ref{thm:chi-squared-exact-boundary} with slowly 
vanishing nominal FWER levels achieve asymptotically exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}. \\

{\rm (ii)} Conversely, if $\overline{r} < \widetilde{g}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, the exact-approximate support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

Our third asymptotic result characterizes the phase transition phenomenon in the approximate support recovery problem in the chi-square model.
It closely parallels Theorem \ref{thm:Gaussian-error-approx-boundary} for the additive errors model.

\begin{theorem} \label{thm:chi-squared-approx-boundary}
Consider the high-dimensional chi-squared model \eqref{eq:model-chisq-Chapter6} with signal sparsity and size as described in \eqref{eq:signal-sparsity} and \eqref{eq:signal-size}.
The function 
\begin{equation} \label{eq:approx-boundary-chisquared}
    h(\beta) = \beta
\end{equation}
characterizes the phase transition of approximate support recovery problem. Specifically the following two results hold.\\

{\rm (i)} If $\underline{r} > {h}(\beta)$, then the \ac{BH} procedure $\widehat{S}_p$ (defined in Section \ref{sec:statistical-procedures}) with slowly vanishing (see Definition \ref{def:slowly-vanishing}) nominal FDR levels achieves asymptotically approximate support recovery in the sense of \eqref{eq:support-recovery-success}. \\

{\rm (ii)} Conversely, if $\overline{r} < {h}(\beta)$, then approximate support recovery asymptotically fails in the sense of \eqref{eq:support-recovery-failure} for all thresholding procedures.
\end{theorem}

@@

\begin{theorem} \label{thm:chi-squared-approx-exact-boundary}
In the context of Theorem \ref{thm:chi-squared-approx-boundary}, the function 
\begin{equation} \label{eq:approx-exact-boundary-chisquared}
    \widetilde{h}(\beta) = \left(\sqrt{\beta} + \sqrt{1-\beta}\right)^2
\end{equation}
characterizes the phase transition of approximate-exact support recovery problem. Namely, 
the following two results hold.\\

{\rm (i)} If $\underline{r} > \widetilde{h}(\beta)$, then the Benjamini-Hochberg procedure with slowly vanishing nominal FDR levels achieves asymptotically approximate-exact support recovery in the sense of \eqref{eq:support-recovery-success}. \\

{\rm (ii)} Conversely, if $\overline{r} < \widetilde{h}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, the approximate-exact support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

@@

Notice that all phase transitions boundaries are identical to those in the Gaussian additive error model \eqref{eq:model-additive} under one-sided alternative.
We refer readers to Figure \ref{fig:phase-Gaussian-errors} in Section \ref{sec:additive-error-model-boundaries} for a visualization of the results in Theorems \ref{thm:chi-squared-exact-boundary} through \ref{thm:chi-squared-approx-exact-boundary}.

@@

All four phase transitions results in Theorems \ref{thm:chi-squared-exact-boundary} through \ref{thm:chi-squared-approx-exact-boundary} 
focus only on the idealized models \eqref{eq:model-chisq-Chapter6} where the statistics are \emph{independent}.
Support recovery problems under dependent observations remain to be explored.
Recall in Chapter \ref{chap:exact-support-recovery} we showed that the boundary for the exact support recovery problem in the additive error model 
\eqref{eq:model-additive} continues to hold even under {\em severe dependence} and general distributional assumptions.
We conjecture that the concentration of maxima phenomenon, which is at the heart of the results in Chapter \ref{chap:exact-support-recovery}, will play
a role and all of the above phase-transition results will continue to hold, under broad dependence conditions in the chi-square models. As an example, 
in the GWAS application, dependence among the genetic markers at different locations (known as linkage disequilibrium) decay as a function of their physical distances on the genome \citep{bush2012genome}, resulting in locally dependent test statistics.
It would be of great interest to extend the current theory to cover important dependence structures that arise in such applications.

@@

% $\mathrm{risk}^{\mathrm{EA}}$ and $\mathrm{risk}^{\mathrm{AE}}$.
As alluded to in Chapter \ref{sec:motivation-chisq} in the introduction, we draw explicit comparisons between the one-sided and two-sided alternatives in Gaussian additive error models \eqref{eq:model-additive}.
% The exact, and the approximate support recovery problems in the additive error model \eqref{eq:model-additive} under standard Gaussian errors have been studied in \cite{gao2018fundamental} and \cite{arias2017distribution}, respectively.

@@

Comparing Theorems \ref{thm:chi-squared-exact-approx-boundary} and \ref{thm:Gaussian-error-exact-approx-boundary} as well as
 Theorems \ref{thm:chi-squared-approx-exact-boundary} and \ref{thm:Gaussian-error-approx-exact-boundary}, we see that the phase transition boundaries under the two types of alternatives are also identical in the exact-approximate and approximate-exact support recovery problems.

@@

Therefore, all phase transition boundaries coincide with those in the additive error models obtained in Chapter \ref{chap:phase-transitions} under their respective parametrizations.
% and Figure \ref{fig:phase-chi-squared} continues to apply.
This indicates vanishing differences between the difficulties of the one-sided and two-sided alternatives in the Gaussian additive error model 
\eqref{eq:model-additive}. The additional uncertainty in the two-sided alternatives does not call for larger signal sizes in these problems, asymptotically.

### 2021-03-10 7.GWAS/7.numerical.tex
ch 7
i implemented my minor edits to ch 7.  I have 3 paragraphs in BLUE which need to be checked -- pages 97, 102, 105.

We illustrate with simulations the phase transition phenomena in the chi-square model, and compare numerically the required signal sizes in support recovery problems between the two types of alternatives in the additive error model.
% We also demonstrate the fundamental trade-off between odds ratios and relative frequencies in association studies, as outlined in Proposition \ref{prop:signal-size-odds-ratio} and Corollary \ref{cor:signal-size-odds-ratio-conditional-frequency}, using evidence from large-scale genetic studies.

\subsection{Exact support recovery}

@@

Experiments were repeated 1000 times for a grid of signal size values ranging from $r=0$ to $6$, and for dimensions $10^2, 10^3$, and $10^5$.
Results of the experiments, shown in Figure \ref{fig:one-vs-two-sided-exact_support_recovery}, suggest vanishing difference between difficulties of two-sided vs one-sided alternatives in the additive error models, as well as vanishing difference between the powers of Bonferroni's procedures and the oracle procedures as $p\to\infty$.

@@

Similar experiments are conducted to examine the optimality claims in Theorem \ref{thm:chi-squared-approx-boundary}, and in Section \ref{subsec:one-vs-two-sided}.
We define an oracle thresholding procedure for approximate support recovery, where the threshold is chosen to minimize the empirical risk.
That is,
$$
t_p(x, S) \in \argmin_{t\in\R} \frac{|\widehat{S}(t)\setminus S|}{\max\{|\widehat{S}(t)|,1\}} + \frac{|S\setminus \widehat{S}(t)|}{\max\{|{S}|,1\}},
%\mathcal{R^\mathrm{oracle}} \in \argmin_{\widehat{S}(\mathcal{R})\in\mathcal{S}} \mathrm{risk}^{\mathrm{A}}(\mathcal{R}),
$$
where $\widehat{S}(t) = \{i\;|\;x(i)\ge t\}$;
in implementation, we only need to scan the values of observations $t\in\{x(1), \ldots, x(p)\}$. 
The nominal FDR level for the BH procedure is set at $1/(5{\log{p}})$, therefore slowly vanishing, in line with the assumptions in Theorem \ref{thm:chi-squared-approx-boundary}; all other parameters are identical to that in the experiments for exact support recovery.
Results of the experiments are shown in Figure \ref{fig:one-vs-two-sided-approx_support_recovery} and in the middle column of Figure \ref{fig:phase-simulated-chi-squared}.

We also examine the boundary described in Theorem \ref{thm:chi-squared-exact-approx-boundary}.
Experimental settings are identical to that in the experiments for approximate support recovery.
% Results for the BH procedure are in general close to that of the oracle procedure,
We compare the performance of the BH procedure with an oracle procedure with threshold
$$
t_p(x, S) \in \min_{i\in S} x(i),
$$
and visualize results of the experiments in the right column of Figure \ref{fig:phase-simulated-chi-squared}.
Notice that the BH procedure sets its threshold somewhat higher than the oracle, especially for small $\beta$'s. 
The empirical risk of the oracle procedure (not shown here in the interest of space) follows much more closely the predicted boundary \eqref{eq:approx-exact-boundary-chisquared}.

>>>>

We illustrate with simulations the phase transition phenomena in the chi-square model, and compare numerically the required signal sizes in support recovery problems between the two types of alternatives in the additive error model.  
\stilian{These results can provide further guidance to practitioners on when to expect genetic association discoveries.
They provide also an explanation of the empirically observed phenomenon known as steep part of the power curve
in statistical genetics \citep{bush2012genome}. }{\fbox{IS this the RIGHT reference?  I could not find the ``steep part of the power curve" anywhere.}}
% We also demonstrate the fundamental trade-off between odds ratios and relative frequencies in association studies, as outlined in Proposition \ref{prop:signal-size-odds-ratio} and Corollary \ref{cor:signal-size-odds-ratio-conditional-frequency}, using evidence from large-scale genetic studies.

\subsection{Exact support recovery} \label{sec:ch7-exact-support-numerics}

@@

The experiments were repeated 1000 times for a grid of signal size values ranging from $r=0$ to $6$, and for dimensions $10^2, 10^3$, and $10^5$.
Results of the experiments, shown in Figure \ref{fig:one-vs-two-sided-exact_support_recovery}, suggest vanishing difference between difficulties of two-sided vs one-sided alternatives in the additive error models, as well as vanishing difference between the powers of Bonferroni's procedures and the oracle procedures as $p\to\infty$.

@@

Similar experiments are conducted to examine the optimality claims in Theorem \ref{thm:chi-squared-approx-boundary}, and in Section \ref{subsec:one-vs-two-sided}.
We define an oracle thresholding procedure for approximate support recovery, where the threshold is chosen to minimize the empirical risk.
That is,
$$
t_p(x, S) \in \argmin_{t\in\R} \frac{|\widehat{S}(t)\setminus S|}{\max\{|\widehat{S}(t)|,1\}} + \frac{|S\setminus \widehat{S}(t)|}{\max\{|{S}|,1\}},
%\mathcal{R^\mathrm{oracle}} \in \argmin_{\widehat{S}(\mathcal{R})\in\mathcal{S}} \mathrm{risk}^{\mathrm{A}}(\mathcal{R}),
$$
where $\widehat{S}(t) = \{i\;|\;x(i)\ge t\}$;
in implementation, we only need to scan the values of observations $t\in\{x(1), \ldots, x(p)\}$. 
The nominal FDR level for the BH procedure is set at $1/(5{\log{p}})$, therefore slowly vanishing, in line with the assumptions in Theorem \ref{thm:chi-squared-approx-boundary}; all other parameters are identical to that in the experiments for exact support recovery in 
Section \ref{sec:ch7-exact-support-numerics}.  The results of the experiments are shown in 
Figure \ref{fig:one-vs-two-sided-approx_support_recovery} and in the middle column of Figure \ref{fig:phase-simulated-chi-squared}.

We also examine the boundary described in Theorem \ref{thm:chi-squared-exact-approx-boundary}.
Experimental settings are identical to that in the experiments for approximate support recovery.
% Results for the BH procedure are in general close to that of the oracle procedure,
We compare the performance of the BH procedure with an oracle procedure with threshold
$$
t_p(x, S) \in \min_{i\in S} x(i),
$$
and visualize the results of the experiments in the right column of Figure \ref{fig:phase-simulated-chi-squared}.
Notice that the BH procedure sets its threshold somewhat higher than the oracle, especially for small $\beta$'s. 
The empirical risk of the oracle procedure (not shown here in the interest of space) follows much more closely the predicted boundary \eqref{eq:approx-exact-boundary-chisquared}.
